---
title: "Regularized Linear Model"
output: html_notebook
---
This is an R Notebook, its comprised of code chunks that you can run in RStudio by pressing Ctrl+Alt+I

```{r}
library(ggplot2)
library(plyr)
library(fBasics)
library(caret)
train <- read.csv("Data/train.csv",stringsAsFactors = FALSE)
test <- read.csv("Data/test.csv", stringsAsFactors = FALSE)
```

```{r}
dim(train)
all_data <- rbind(train[,c(1:80)],test)
dim(all_data)
```
Creates a data frame that combines the test and train data
```{r}
df <- rbind(data.frame(version="log(price+1)",x=log(train$SalePrice + 1)),
            data.frame(version="price",x=train$SalePrice))
ggplot(data=df) +
  facet_wrap(~version,ncol=2,scales="free_x") +
  geom_histogram(aes(x=x))
train$SalePrice <- log(train$SalePrice + 1)
```
See the before and after of transforming the sale prices to remove skewness

Combining BsmtFinSF1 and BsmtFinSF2 variables
```{r}
#all_data$BsmtFinSF1 <- all_data$BsmtFinSF1 + all_data$BsmtFinSF2
#colnames(BsmtFinSF1) <- "BsmtFinSF"
```

Removing least useful predictors
```{r}
all_data <- subset(all_data, select=-c(Street, Utilities, PoolArea, MiscFeature))
```

Numerical encoding ordered categorical features, like quality
```{r}
all_data$ExterQual[all_data$ExterQual == 'Ex'] <- 4
all_data$ExterQual[all_data$ExterQual == 'Gd'] <- 3
all_data$ExterQual[all_data$ExterQual == 'TA'] <- 2
all_data$ExterQual[all_data$ExterQual == 'Fa'] <- 1
all_data$ExterQual[all_data$ExterQual == 'Po'] <- 0

all_data$ExterCond[all_data$ExterCond == 'Ex'] <- 4
all_data$ExterCond[all_data$ExterCond == 'Gd'] <- 3
all_data$ExterCond[all_data$ExterCond == 'TA'] <- 2
all_data$ExterCond[all_data$ExterCond == 'Fa'] <- 1
all_data$ExterCond[all_data$ExterCond == 'Po'] <- 0

all_data$HeatingQC[all_data$HeatingQC == 'Ex'] <- 4
all_data$HeatingQC[all_data$HeatingQC == 'Gd'] <- 3
all_data$HeatingQC[all_data$HeatingQC == 'TA'] <- 2
all_data$HeatingQC[all_data$HeatingQC == 'Fa'] <- 1
all_data$HeatingQC[all_data$HeatingQC == 'Po'] <- 0

all_data$KitchenQual[all_data$KitchenQual == 'Ex'] <- 4
all_data$KitchenQual[all_data$KitchenQual == 'Gd'] <- 3
all_data$KitchenQual[all_data$KitchenQual == 'TA'] <- 2
all_data$KitchenQual[all_data$KitchenQual == 'Fa'] <- 1
all_data$KitchenQual[all_data$KitchenQual == 'Po'] <- 0

all_data$Functional[all_data$Functional == 'Typ'] <- 7
all_data$Functional[all_data$Functional == 'Min1'] <- 6
all_data$Functional[all_data$Functional == 'Min2'] <- 5
all_data$Functional[all_data$Functional == 'Mod'] <- 4
all_data$Functional[all_data$Functional == 'Maj1'] <- 3
all_data$Functional[all_data$Functional == 'Maj2'] <- 2
all_data$Functional[all_data$Functional == 'Sev'] <- 1
all_data$Functional[all_data$Functional == 'Sal'] <- 0

all_data$FireplaceQu[all_data$FireplaceQu == 'Ex'] <- 5
all_data$FireplaceQu[all_data$FireplaceQu == 'Gd'] <- 4
all_data$FireplaceQu[all_data$FireplaceQu == 'TA'] <- 3
all_data$FireplaceQu[all_data$FireplaceQu == 'Fa'] <- 2
all_data$FireplaceQu[all_data$FireplaceQu == 'Po'] <- 1
all_data$FireplaceQu[all_data$FireplaceQu == 'NA'] <- 0

all_data$GarageFinish[all_data$GarageFinish == 'Fin'] <- 3
all_data$GarageFinish[all_data$GarageFinish == 'RFn'] <- 2
all_data$GarageFinish[all_data$GarageFinish == 'Unf'] <- 1
all_data$GarageFinish[all_data$GarageFinish == 'NA'] <- 0

all_data$GarageQual[all_data$GarageQual == 'Ex'] <- 5
all_data$GarageQual[all_data$GarageQual == 'Gd'] <- 4
all_data$GarageQual[all_data$GarageQual == 'TA'] <- 3
all_data$GarageQual[all_data$GarageQual == 'Fa'] <- 2
all_data$GarageQual[all_data$GarageQual == 'Po'] <- 1
all_data$GarageQual[all_data$GarageQual == 'NA'] <- 0

all_data$GarageCond[all_data$GarageCond == 'Ex'] <- 5
all_data$GarageCond[all_data$GarageCond == 'Gd'] <- 4
all_data$GarageCond[all_data$GarageCond == 'TA'] <- 3
all_data$GarageCond[all_data$GarageCond == 'Fa'] <- 2
all_data$GarageCond[all_data$GarageCond == 'Po'] <- 1
all_data$GarageCond[all_data$GarageCond == 'NA'] <- 0

all_data$PoolQC[all_data$PoolQC == 'Ex'] <- 4
all_data$PoolQC[all_data$PoolQC == 'Gd'] <- 3
all_data$PoolQC[all_data$PoolQC == 'TA'] <- 2
all_data$PoolQC[all_data$PoolQC == 'Fa'] <- 1
all_data$PoolQC[all_data$PoolQC == 'NA'] <- 0

all_data$Fence[all_data$Fence == 'GdPrv'] <- 4
all_data$Fence[all_data$Fence == 'MnPrv'] <- 3
all_data$Fence[all_data$Fence == 'GdWo'] <- 2
all_data$Fence[all_data$Fence == 'MnWw'] <- 1
all_data$Fence[all_data$Fence == 'NA'] <- 0
```

```{r}
feature_classes <- sapply(names(all_data),function(x){class(all_data[[x]])})
```

sapply applies the second argument, a function that returns the type of feature each column belongs to() either numeric or character). feature_classes is an array that contains feature types, and the indices correspond to the column numbers in the data
```{r}
numeric_feats <-names(feature_classes[feature_classes != "character"])
# determine skew for each numeric feature
skewed_feats <- sapply(numeric_feats,function(x){skewness(all_data[[x]],na.rm=TRUE)})
# keep only features that exceed a threshold for skewness
skewed_feats <- skewed_feats[skewed_feats > 0.75]

# transform excessively skewed features with log(x + 1)
for(x in names(skewed_feats)) {
  all_data[[x]] <- log(all_data[[x]] + 1)
}
```

```{r}
# get names of categorical features
categorical_feats <- names(feature_classes[feature_classes == "character"])
#remove features that have already been encoded above
remove <- c('ExterQual','ExterCond', 'HeatingQC', 'KitchenQual', 'Functional', 'FireplaceQu', 'GarageFinish', 'GarageQual', 'GarageCond', 'PoolQC', 'Fence')
categorical_feats <- categorical_feats[! categorical_feats %in% remove]
# use dummyVars function from caret library for dummy encoding
dummies <- dummyVars(~.,all_data[categorical_feats])
categorical_1_hot <- predict(dummies,all_data[categorical_feats])
categorical_1_hot[is.na(categorical_1_hot)] <- 0  #for any level that was NA, set to zero
```

```{r}
# for any missing values in numeric features, impute mean of that feature
numeric_df <- all_data[numeric_feats]

for (x in numeric_feats) {
    mean_value <- mean(train[[x]],na.rm = TRUE)
    all_data[[x]][is.na(all_data[[x]])] <- mean_value
}
```

```{r}
a = c(0,1,1,1,0, 0)
b = c(1,0,1,0,1, 0)
(a+b)%%2
```

```{r}
# reconstruct all_data with pre-processed data
all_data <- cbind(all_data[numeric_feats],categorical_1_hot)
#remove ElectricalMix category
all_data <- subset(all_data, select=-c(ElectricalMix))
all_data[,1] <- NULL

#combine the Ext1 and Ext2 variables with element-wise logical OR
combine_ext <- matrix(FALSE, nrow = 2919, ncol=16)
for (i in 1:15){
  combine_ext[,i] <- all_data[127+i] | all_data[144+i]
}
combine_ext <- combine_ext+0
all_data[,128:143] <- combine_ext
all_data[,143:158] <- NULL
```
Creating a new variable BsmtQual
```{r}
exp(12)
```

```{r}
# create data for training and test
X_train <- all_data[1:nrow(train),]
X_test <- all_data[(nrow(train)+1):nrow(all_data),]
y <- train$SalePrice
all_train <-  cbind(X_train, y)

write.csv(all_train, file = 'Data/trainy_processed.csv')
write.csv(X_test, file = 'Data/test_processed.csv')
```

